Efficient Coalition Formation for Web Services

### 2013-03-16 papers/coalition/coalition.tex
Changes

\usepackage{amssymb}
\usepackage{amsthm}

\theoremstyle{plain}
\newtheorem{thm}{Theorem} % reset theorem numbering for each chapter

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition} % definition numbers are dependent on theorem numbers
\newtheorem{exmp}[thm]{Example} % same for example numbers
% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

@@

\author{\IEEEauthorblockN{Ehsan Khosrowshahi Asl, Jamal Bentahar}
\IEEEauthorblockA{Faculty of Engineeing and Computer Science\\
Concordia University\\
Montreal, Canada\\
e\_khosr@encs.concordia.ca, bentahar@ciise.concordia.ca}
\and
\IEEEauthorblockN{Hadi Otrok, Rabeb Mizouni}
\IEEEauthorblockA{Department of Electrical and Computer Engineering\\
Khalifa University\\
Dubai, UAE\\
hadi.otrak, rabeb.mizouni@kustar:ac:ae}
}

@@

Web services are loosely-coupled business applications and willing to cooperate in distributed settings within different groups called communities. Communities aim to provide better efficiency, market share and total payoff. There are a number of proposed mechanisms and models on aggregating web services, however forming optimal and stable coalitions to maximize individual and community efficiency and income has got less attention. In this paper, we propose an efficient coalition formation mechanism using cooperative game-theoretic methods. We propose a mechanism for membership requests and selection of web services in the scenarios in which there are already established communities. Moreover, we analyze the scenarios in which we do not have pre-defined communities and we develop a mechanism for web services to form stable groups allowing them to maximize their efficiency and generate near-optimal (welfare-maximizing) communities. The theoretical and simulation results show that our algorithms provide web services and community owners with applicable and near-optimal decision making mechanisms.

@@

\section{Introduction}

Over the past years, online services have become part of critical business services. The increasing reliance on web-based applications has significant implications for service providers which calls for a seamless integration of business process, applications and web services of the Internet. Web services as a highly available software function accessible at a network address over the web, usually provide a set of stateless functions. Delivering high quality services on the dynamic and unpredictable nature of Internet is a critical and challenging task.

The need for highly available and highly responsive services online has called for collaborative communities of these loosely-coupled business web services. The idea of web service communities and the way they are engineered to collaborate was proposed in \cite{DBLP:journals/ijebr/MaamarSTBB09, DBLP:journals/internet/BenatallahSD03, Rosario:2008:PQS:1512146.1512290} where web services with similar functionality are grouped together and users send their requests to communities assuring higher service availability and performance than individual web services. These communities are highly available also more resilience to failure since there is no single point of failure in their architecture.

Most of the recent work on communities of services are either user-centric and focus on user satisfaction \cite{Chun02user-centricperformance} or they are system-centric and focus on system throughput and performance and utilization of the system. There are many contributions in distributed, grid, cluster and cloud computing services which are system-centric. However in real world environments both users and service providers are self-interested agents, aiming to maximize their own profit. In real world environments, both these parties will participate and collaborate as long as it's economic and efficient for both of them.

Recently, \cite{DBLP:conf/IEEEscc/LimTMB12, DBLP:conf/IEEEscc/KhosravifarABT11} have proposed mechanisms in which they try to maximize gain for both parties. A two-player non-cooperative game theory method between web services and community manager was proposed in \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. The strategies for a web service when facing a new community, is requesting to join or refusing the invitation to join the community and the set of strategies for communities are inviting the web service or refusing the web service join request. Both parties based on their capacity, market share and reputation will have different set of utilities over the strategy profiles of the game. The problem suffers from ignoring many quality parameters and also ignoring the web services already residing within the community. The game is only between the community master and the new web service, and all other web services already residing within the community are ignored. Those web services can lose utility or payoff, having to collaborate with new web services. They may deviate and join other communities if they are unsatisfied, which is not addressed in their approach. In \cite{DBLP:conf/IEEEscc/LimTMB12} a 3-Way satisfaction selection approach has been proposed. In their approach, they have proposed a web service selection process for community masters catering for the efficiency of all three parties of users, web services and communities. In this work, they show how the gains of these parties are coupled together and they suggest a linear optimization process for optimizing the needs of the 3 parties. However, the optimization problem in this solution tends to optimize some parameters considering all web services even if they are weak. There are no clear thresholds for accepting or rejecting new web services. The optimization result for example could suggest web services already residing within the community to increase or decrease their capacity to adjust with weakness of other parties in the system. However, a strong performing web service could deviate anytime it finds itself unsatisfied within the community instead of adjusting its service parameters.

In this paper, we propose a cooperative game theory based model for aggregation of web services within communities. The solution concepts of cooperative game theory seeks to find most efficient ways of forming coalitions (teams) of agents, maximizing their gain and payoff, and also distributing gain in a fair way to keep coalitions stable by having all agents satisfied. We first propose a representation function for communities of web services based on their qualities and capacities. Using this function we can evaluate the $worth$ of each community of web services. Our community managers facing new membership requests, analyses whether the new coalition having the new set of web services will keep the community stable for $all$ web services already residing in the community or not. The community manager will reject the membership requests if it finds out new coalition would be unstable, preventing $any$ subset of web services from gaining significantly more, by deviating and joining other communities. Also web services facing communities, will try to join communities which maximizes their gain, if they find out they can deviate and gain more, they will form new coalitions with other communities. The cooperative game theory solution methods are combinatorial in nature and are NP-complete, making them impractical in real world applications; however, using the concepts of coalition stability we have offered approximation algorithms running in polynomial time providing web services and community managers with applicable and near-optimal decision making mechanisms.

The rest of paper is organized as follows. Section \ref{s:preliminaries} describes architecture and parameters and some basic cooperative game theory concepts, section \ref{s:model} presents our solution model in three different scenarios. Section \ref{s:resutls} presents our experiments and results and finally, section \ref{s:conclusion}, concludes the paper and identified future work.

@@

In this section we discuss the parameters and preliminary concepts that we use in the rest of paper.

@@

Our system consists of three main type of entities working together:

\emph{1) Web services} are rational entities providing services to end users. They aim to maximize their individual income by receiving enough requests from end users. In order to increase their revenue, web services seek for more tasks if they have the capacity and throughput. Web services can join communities to have better efficiency by collaborating with others, to have access to higher market share and to have opportunity of receiving a bigger task pool from end users. The way the web services reside inside communities and how communities of web services are engineered is described thoroughly in \cite{DBLP:journals/ijebr/MaamarSTBB09}. Throughout this paper, in our equations, we refer to web services as $ws$.

\emph{2) Master Web Services} are responsible for managing and are representative of the communities of web services. Communities receieve tasks from users and they aim to have a healthy set of web services to perform the tasks. They seek to maximise user satisfaction by having tasks accomplished with good rate and quality. Higher user satisfaction will bring more user tasks and increase the market share and revenue of the community.

\emph{3) Users} generate tasks and try to find service providers online to perform the tasks. User satisfaction is abstracted as function of quantity and quality of tasks accomplished by our service providers.

\subsection{Web Sercive Parameters}

@@

\begin{table}[!t]
\centering
\caption{List of web service QoS parameters.}
\begin{tabular}{|c|c||c|c|}
\hline
\textbf{Notation} & \textbf{Definition} \\
\hline\hline
$Availability$ & Probability of available during a time frame \\
$Reliability$ & P of successfully handling tasks during a timeframe\\
$Successability$ & Probability of handling tasks successfully \\
$Throughput$ & Average rate of handling tasks \\
$Latency$ & The average latency of services\\
$Capacity$ & Amount of resources available\\
$Cost$ & Mean service fee for cooperating \\
$Regulatory$ & Compliance with standards, law and rules\\
$Security$ & Quality of confidentiality and non-repudiation\\
\hline
\end{tabular}
\label{qosws}
\end{table}

We adopted a real world dataset \cite{DBLP:conf/smc/Al-MasriM09a} which has aggregated and normalized each of these parameters to a real value between 0 and 1. Since tasks are not shared between web services and are distributed among all web services inside a community, one can expect tasks within a community to be performed by average quality of web service residing in that community. We use this expected quality output $(QoS_{ws})$, later in evaluating community \emph{worth} or \emph{payoff} function.

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies strategies of self-interested agents in a setting where agents can increase their payoff by binding agreements and cooperating in groups. We let $N$ to be a set of players. Any subset $S$ of $N$ can form a group that we call it a $coalition$. A \emph{coalitional game} is a pair $G = (N, v)$ where $v$ is called a \emph{characteristic function}, which given a coalition, is a function $v: 2^N \to \mathbb{R}$ mapping the set of players of the coalition to a real number $v(S)$, the worth of $S$. This number usually resembles the output or payoff or performance of these players working together as a coalition.  If a coalition $S$ forms then it can divide its worth, $v(S)$ in any possible way among its members. The payoff vector $x \in \mathbb{R}^S$ is the amount of payoff being distributed among agents of coalition $S$. The payoff vector satisfies two conditions:

\begin{itemize}
	\item $x \geq 0$ for all $i \in N$, and
	\item $\sum_{i \in S} x_i \leq v(S)$
\end{itemize}

The second criteria is the \emph{feasibility} condition, according to the equation, the payoff for each agent cannot be more than coalitions total gain. A payoff vector is also \emph{efficient} if all the payoff obtained by a coalition is distributed amongst coalition members: $\sum_{i \in S} x_i = v(S)$. This definition of characteristic function works in \emph{transferable utility} (TU) settings, where utility (payoff) is transferable from one player to other, in other words players have common currency and a unit of income is worth same for all players \cite{myerson1991game}.

When dealing with cooperative games, two issues need to get addressed: 1. What coalition to form (Canonical games)? 2. How to reward each member when a task is completed (Coalition-formation games)?

@@

{\bf Definition 1 (Shapley value)} Given a cooperative game (N, v), the \emph{Shapley value} of player $i$ is given by\cite{shapley_value}: 
\begin{equation}\label{eq:shapley}
\phi_i(v) = \sum_{S \subseteq N \backslash \left\{i\right\} } \frac{|S|! (n-|S|-1)!}{n!} (v(S \cup \left\{i\right\}) - v(S))
\end{equation}

\emph{Shapley value} is a unique and fair solution concept for payoff distribution between agents. It basically rewards agents with the amount of marginal contribution they have to the coalition.

@@

{\bf Definition 2 (Core)} A payoff vector $\mathbf{x}$, denoted by $C(N,v)$, is in the $core$ of a coalitional game $(N, v)$ if and only if
\begin{equation}\label{eq:core}
\forall S \subseteq N, \sum_{x_i \in S} x_i \geq v(S)
\end{equation}

The core is basically a set of payoff vectors $(\mathbf{x})$ where no subset of players $S^\prime$ could gain more than their current payoff, by deviating and making their own coalition $\sum_{i \in S^\prime} x_i \geq v(S^\prime)$. The sum of payoffs to the players in any sub-coalition $S$ is at least as large as the amount that these players could earn by forming a coalition on their own. In a sense it is analogue to Nash equilibrium, except that it about deviations by groups of agents. The core is the strongest and most popular solution concept in cooperative game theory, however it is a combinatorial problem and is not computationally tractable as number of players increase. The core of some real-world problem games may be empty, which means having characteristic function game of $(N,v)$, there might be no possible distribution of payoff assuring stability of subgroups.

{\bf Definition 3 (Convex cooperative games)} A game with characteristic function $v(S)$ is convex if:
\begin{equation}\label{eq:convex}
V(S) + V(T) \leq v(S \cup T) + v (S \cap T), \forall S,T \subseteq N.
\end{equation}

@@

\subsubsection*{$\epsilon$-core}\label{s:epsilon}
%\emph{$\epsilon$-Core:}
%\\
When \emph{core} set of a game is empty, it means no coalition of players can gain anything by deviating. An outcome would be unstable if a coalition can benefit even by a small amount from deviating. This is a strong requirement. In some situations, deviations can be costly or players may have loyalty to their coalitions, or even it can be computationally intractable to find those small benefits. It would only make sense for a coalition to deviate if the gain from a deviation exceeded a particular cost for example the cost of deviation. \emph{$\epsilon$-core} relaxes the notion of the core, and only requires that, no coalition would benefit significantly, or within a constant amount($\epsilon$) by deviating.

@@

Coalition structure formation is the problem of finding best partition of agents into teams. In these settings, performance of individual agent is less important than the \emph{social welfare} of the system, which is the sum of the values of all teams. Having characteristic function game $G(N,v)$, a coalition structure $(CS)$ is \emph{socially optimal} if $CS$ belongs to set $argmax_{CS \in CS_N}v(CS)$ where $v(CS)$ is sum of the values of all coalitions inside $CS$. $v(CS) = \sum_{C \in CS_N}v(CS)$. 
%The outcome of a characteristic function game in coalition structure settings, consists of two parts; first a disjoint partition of players (agents) into coalitions, called a \emph{coalition structure} (CS) and second a \emph{payoff vector} as mentioned in cooperative game solution concepts, which distributes the value of each coalition among its members.

@@

In this section, we present the architecture of different web service community models and formulate how both web services and community masters as rational agents, would adopt strategies to maximize their payoff.

\subsection {Web Services and A Grand Community}

In this scenario, there is a community controller by its master web service. The community master has a task request rate $(R_{master})$ from users. Web services need to join a community to be able to get tasks from a master web service. Each web service comes with different quality of service parameters and a throughput $(T_{ws})$. Throughput is the average rate of tasks a web service can perform per time unit. Therefore, the master web service is providing tasks with rate of $R_{ws}$ and web services perform tasks with average output quality of $QoS_{ws}$ and a rate or $T_{ws}$. 
In this setting, we define the value of the coalition of the web services within a community as follows:
\begin{equation*}
output(C) = \sum_{ws}{(T_{ws} \times QoS_{ws})}
\end{equation*}
\setlength{\abovedisplayshortskip}{2pt}
\begin{equation*}
v(C) = \left\{ 
  \begin{array}{l l}
    output(C) & \quad \text{if $\sum_{ws}{T_{ws}} \leq R_{master}$}\\
    output(C) \times \frac{R_{master}}{\sum_{ws}{T_{ws}}} & \quad \text{if $\sum_{ws}{T_{ws}} > R_{master}$}
  \end{array} \right.
\end{equation*}

The output of a coalition of web services is the rate and quality of task they can perform. If the rate is more than master's input task rate, it means the web services inside the community are capable of doing more tasks than exists. Considering this the valuation function will balance the output performance and will be the exact rate and quality of work they can perform within the particular community. %** In the case where the limited tasks are distributed among web services uniformly, the value of coalition would be the proportion of the average QoS times their throughput to rate of available requests. **

In the first scenario, we only consider one grand coalition and analyse the system from point of view of one single master web service and a collection of web services. The master web service decides which members can join %or should leave% 
the community and distributes the income or tasks among its community members.

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{s1.eps}`
\caption{Web Services and A Grand Community}
\label{fig_sim}
\end{figure}

The membership decision is made based on throughput and \emph{QoS} parameters of the requested web service. The goal is to have quality web services in community so the community stays stable and no other web services would have incentives to deviate and leave the coalitions. Therefore, a basic method would be to check the core membership of the coalition of all current community members (all web services already residing within the community) and the new web service. This algorithm would use the \emph{Shapley value} distribution method as described in equation \ref{eq:shapley} to distribute gain of $v(C)$ among all members and then checks if \emph{Shapley value} payoff vector for this community having characteristic function $v(C)$ is \emph{core} stable or not. In \emph{Shapley value} payoff vector, the payoff for each web service $ws_i$, is calculated based on their marginal contribution $v(C \cup {i}) - v(C)$ over all the possible different permutations in which the coalition can be formed, this is why \emph{Shapley value} distribution is fair. Because of going through all the possible different permutations of subsets of $N$, the nature of the \emph{Shapley value} is combinatorial. This makes it impractical to use as size of our coalitions grow.

It is proven, in convex games the \emph{Shapley value} lies in the core \cite{DBLP:conf/ijcai/GrecoMPS11, myerson1991game}. Therefore if the \emph{Core} is non-empty, the \emph{Shapley value} payoff vector is a member of \emph{Core}.
% so in our algorithm we check the core membership of this payoff vector.

@@

This implies in order to keep the characteristic function convex, new agents should have more marginal contribution as coalition size grows.

>>>>

\usepackage{amssymb}
\usepackage{amsthm}
\pagestyle{plain}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}% reset theorem numbering for each chapter
\newtheorem{Proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition} % definition numbers are dependent on theorem numbers
\newtheorem{exmp}[thm]{Example} % same for example numbers
% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

@@

\author{\IEEEauthorblockN{Ehsan Khosrowshahi Asl, Jamal Bentahar}
\IEEEauthorblockA{Faculty of Engineeing and Computer Science\\
Concordia University\\
Montreal, Canada\\
e\_khosr@encs.concordia.ca, bentahar@ciise.concordia.ca}
\and
\IEEEauthorblockN{Hadi Otrok, Rabeb Mizouni}
\IEEEauthorblockA{Department of Electrical and Computer Engineering\\
Khalifa University of Science, Technology and Research\\
Abu Dhabi, UAE\\
hadi.otrak, rabeb.mizouni@kustar:ac:ae}
}

@@

Web services are loosely-coupled business applications willing to
cooperate in distributed settings within different groups called
communities. Communities aim to provide better visibility,
efficiency, market share and total payoff. There are a number of
proposed mechanisms and models on aggregating web services and
making them cooperate within their communities. However, forming
optimal and stable communities as coalitions to maximize
individual and group efficiency and income has not been addressed
yet. In this paper, we propose an efficient coalition formation
mechanism using cooperative game-theoretic techniques. We propose
a mechanism for community membership requests and selections of
web services in the scenarios where established communities
already exist. Moreover, we analyze the scenarios where
communities are not established yet. The ultimate objective is to
develop a mechanism for web services to form stable groups
allowing them to maximize their efficiency and generate
near-optimal (welfare-maximizing) communities. The theoretical and
simulation results show that our algorithms provide web services
and community owners with applicable and near-optimal decision
making mechanisms.

@@

\section{Introduction and Related Work}

Over the past years, online services have become part of many
scalable business applications. The increasing reliance on
web-based applications has significantly influenced the way web
services are engineered. Web services provide a set of stateless
software functions accessible at a network address over the web.
The recent developments are shifting web services from passive and
individual components to autonomous and group-based components
where interaction, composition, and cooperation are the key
challenges \cite{ICWS2011-1,SCC2011-1}. The main objective is to
achieve a seamless integration of business processes, applications
and web services. Delivering high quality services considering the
dynamic and unpredictable nature of the Internet is still a very
critical and challenging issue.

The need for highly available and responsive services has called
for grouping and collaborative mechanisms of loosely-coupled web
services, particularly in business settings. The idea of grouping
web services within communities and the way those communities are
engineered so that web services can better collaborate have been
proposed and investigated in
\cite{DBLP:journals/ijebr/MaamarSTBB09,DBLP:journals/internet/BenatallahSD03,Rosario:2008:PQS:1512146.1512290}.
Communities are virtual groups of web services having similar
functionalities, but probably different non-functional quality
attributes, which form the QoS parameters. When communities are
used, users send their requests to the masters of those
communities, which are responsible of managing the communities,
forwarding the requests to the suitable member web services and
checking the credentials of those members. Communities aim to
provide higher service availability and performance than what
individual web services can provide. The high availability of
services and the community resilience to failure are guaranteed
since web services can cooperate and replace each other within the
same community and since there is no single point of failure in
the communities architecture.

Most of the recent work on communities of services are either
user-centric and focus on user satisfaction
\cite{Chun02user-centricperformance} or system-centric and focus
on the whole system throughput, performance and utilization. There
are many contributions in distributed, grid, cluster and cloud
services which are system-centric. However, in real world
environments and applications, both users and service providers
are self-interested agents, aiming to maximize their own profit.
In those environments, both parties (users and services) will
collaborate as long as they are getting more benefits and payoff.

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11} proposed mechanisms to help
users and services maximize their gain. A two-player
non-cooperative game between web services and community master was
introduced in \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this
game-theoretic model, the strategies available to a web service
when facing a new community are requesting to join the community,
accepting the master's invitation to join the community, or
refusing the invitation to join. The set of strategies for
communities are inviting the web service or refusing the web
service's join request. Based on their capacity, market share and
reputation, the two players have different set of utilities over
the strategy profiles of the game. The main limits of this game
model are: 1) its consideration of only three quality parameters,
while the other factors are simply ignored; and 2) the
non-consideration of the web services already residing within the
community. The game is only between the community master and the
new web service, and the inputs from all the other members are
simply ignored. The consideration of those inputs is a significant
issue as existing web services can lose utility or payoff because
of the new member, which can results in an unhealthy and unstable
group. The problem comes from the fact that the existing members
should collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

In \cite{DBLP:conf/IEEEscc/LimTMB12} a 3-way satisfaction approach
for selecting web services has been proposed. In this approach,
the authors proposed a web service selection process that the
community masters can use. The approach considerers the efficiency
of all the three involved parties, namely users, web services and
communities. In this work, it is shown how the gains of these
parties are coupled together using a linear optimization process.
However, the optimization problem in this solution tends to
optimize some parameters considering all web services regardless
of their efficiency and contribution to the community's welfare.
Moreover, there are no clear thresholds for accepting or rejecting
new web services. The solution of the optimization problem could,
for instance, suggest web services already residing within the
community to increase or decrease their capacity to cover up the
weakness of other parties in the system. However, a high
performing web service could deviate anytime it finds itself
unsatisfied within the community instead of adjusting its service
parameters.

\textbf{Contributions.} In this paper, we use game theory to
propose a cooperative game model for the aggregation of web
services within communities. The solution concepts of our
cooperative game seeks to find efficient ways of forming
coalitions (teams) of web services so that they can maximize their
gain and payoff, and distribute the gain in a fair way among all
the web services. Achieving Fairness when the gain is distributed
among the community members is the main factor to keep the
coalition stable as no web service will expect to gain better by
deviating for the community. In other words, the coalition is made
efficient if all the members are satisfied. We first propose a
representation function for communities of web services based on
their QoS attributes. By using this function, we can evaluate the
$worth$ of each community of web services. When facing new
membership requests, a typical community master checks whether the
new coalition having the old and new set of web services will keep
the community stable or not. The community master will reject the
membership requests if it finds out that the new coalition would
be unstable, preventing $any$ subset of web services from gaining
significantly more by deviating from the community and joining
other communities or forming new ones. The computation of
solutions for cooperative game theory problems is combinatorial in
nature and proven to be NP-complete \cite{Algorithmic}, making
this computation impractical in real world applications. However,
using the concepts of coalition stability, we propose
approximation algorithms running in polynomial time providing web
services and community masters with applicable and near-optimal
decision making mechanisms.

The rest of paper is organized as follows. Section
\ref{s:preliminaries} describes the architecture, considered
parameters and some basic cooperative game theory concepts.
Section \ref{s:model} presents our solution model in three
different scenarios. Section \ref{s:resutls} describes our
experiments and results. Finally, Section \ref{s:conclusion},
concludes the paper and identifies future work.

@@

In this section, we discuss the parameters and preliminary
concepts that we use in the rest of the paper.

@@

Our system consists of three main types of entities working
together:

\emph{1) Web services} are rational entities\footnote{The term
rational is used here in the sense that web services are utility
maximizers} providing services to end users. They aim to maximize
their individual income by receiving enough requests from end
users. In order to increase their revenue, web services seek for
more tasks if they have the capacity and throughput to do so. Web
services can join communities to have better efficiency by
collaborating with others, to have access to higher market share,
and to have opportunity of receiving a bigger task pool from end
users. The way the web services reside inside communities and how
communities of web services are engineered is described in
\cite{DBLP:journals/ijebr/MaamarSTBB09}. Throughout this paper, in
our equations, we refer to web services as $ws$ and to the set of
web services hosted by a given community as $C$. To simply the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

\emph{2) Master Web Services} are representatives of the
communities of web services and responsible for their management.
Communities receive requests from users and aim to host a healthy
set of web services to perform the required tasks. They seek to
maximize user satisfaction by having tasks accomplished according
to the desired QoS. In fact, higher user satisfaction will bring
more user requests and increase the market share and revenue of
the community.

\emph{3) Users} generate requests and try to find the best
available services. User satisfaction is abstracted as function of
quantity and quality of tasks accomplished by a given service.

\subsection{Web Service Parameters}

@@

\begin{table}[!t]
\centering
\caption{List of web service QoS parameters.}
\begin{tabular}{|c|c||c|c|}
\hline
\textbf{Parameter} & \textbf{Definition} \\
\hline\hline
$Availability$ & Probability of being available during a time frame \\
$Reliability$ & Probability of successfully handling requests \\
&during a timeframe\\
$Successability$ & Rate of successfully handled requests \\
$Throughput$ & Average rate of handling requests \\
$Latency$ & The average latency of services\\
$Capacity$ & Amount of resources available\\
$Cost$ & Mean service fee \\
$Regulatory$ & Compliance with standards, law and rules\\
$Security$ & Quality of confidentiality and non-repudiation\\
\hline
\end{tabular}
\label{qosws}
\end{table}

We adopted a real world dataset \cite{DBLP:conf/smc/Al-MasriM09a}
which has aggregated and normalized each of these parameters to a
real value between 0 and 1. Since requests are not shared among
web services and are distributed among all of them inside a
community, each one of them comes with a given QoS denoted by
$(QoS_{ws})$. We assume that $(QoS_{ws})$ is obtained by a certain
aggregation function of the parameters considered in Table
\ref{qosws}. We use this quality output later in evaluating the
community \emph{worth} or \emph{payoff} function.

\subsection{Cooperative Game Concepts}
Cooperative game is a branch of game theory that studies
strategies of self-interested entities or agents in a setting
where those agents can increase their payoff by binding agreements
and cooperating in groups. We let $N$ be a set of players. Any
subset $S$ of $N$ can form a group called $coalition$. A
\emph{coalitional game} is a pair $G = (N, v)$ where $v$ is called
a \emph{characteristic function}, which given a coalition, is a
function $v: 2^N \to \mathbb{R}$ mapping the set of players of the
coalition to a real number $v(S)$, the worth of $S$. This number
usually represents the output or payoff or again the performance
of these players working together as coalition.  If a coalition
$S$ is formed, then it can divide its worth, $v(S)$ in any
possible way among its members. The payoff vector $x \in
\mathbb{R}^S$ is the amount of payoff being distributed among the
members of the coalition $S$. The payoff vector satisfies two
conditions:

\begin{itemize}
    \item $x_i \geq 0$ for all $i \in N$, and
    \item $\sum_{i \in S} x_i \leq v(S)$
\end{itemize}

The second criteria is called the \emph{feasibility} condition,
according to which, the payoff for each agent cannot be more than
the coalition total gain. A payoff vector is also \emph{efficient}
if the payoff obtained by a coalition is distributed amongst the
coalition members: $\sum_{i \in S} x_i = v(S)$. This definition of
the characteristic function works in \emph{transferable utility}
(TU) settings, where utility (i.e., payoff) is transferable from
one player to another, or in other words, players have common
currency and a unit of income is worth same for all players
\cite{myerson1991game}.

When dealing with cooperative games, two issues need to be
addressed:\\ 1. What coalition to form (Canonical games)? \\
2. How to reward each member when a task is completed
(Coalition-formation games)?\\
%
The following definitions help address these two issues.

@@

{\bf Definition 1 (Shapley value)} Given a cooperative game $(N,
v)$, the \emph{Shapley value} of player $i$ is given
by\cite{shapley_value}:
\begin{equation}\label{eq:shapley}
\phi_i(N,v) = \sum_{S \subseteq N \backslash \left\{i\right\} }
\frac{|S|! (|N|-|S|-1)!}{|N|!} (v(S \cup \left\{i\right\}) - v(S))
\end{equation}

\emph{Shapley value} is a unique and fair solution concept for
payoff distribution among the members of the coalition. It
basically rewards members with the amount of marginal contribution
they have to the coalition.

@@

{\bf Definition 2 (Core)} A payoff vector $x$ is in the $core$ of
a coalitional game $(N, v)$ if and only if:
\begin{equation}\label{eq:core}
\forall S \subseteq N, \sum_{x_i \in S} x_i \geq v(S)
\end{equation}

The core is basically a set of payoff vectors where no subset of
players $S^\prime$ could gain more than their current payoff by
deviating and making their own coalition $\sum_{i \in S^\prime}
x_i \geq v(S^\prime)$. The sum of payoffs of the players in any
sub-coalition $S$ is at least as large as the amount that these
players could earn by forming a coalition by their own. In a
sense, it is analogue to Nash equilibrium, except that core is
about deviations from groups of entities. The core is the
strongest and most popular solution concept in cooperative game
theory. However, its computation is a combinatorial problem and
becomes intractable as the number of players increases. The core
of some real-world problem games may be empty, which means having
the characteristic function of the game $(N,v)$, there might be no
possible distribution of payoff assuring stability of subgroups.

{\bf Definition 3 (Convex cooperative games)} A game $(N,v)$ with
characteristic function $v(S)$ is convex if:
\begin{equation}\label{eq:convex}
v(S) + v(T) \leq v(S \cup T) + v (S \cap T), \forall S,T \subseteq
N.
\end{equation}

@@

\subsubsection*{$\epsilon$-core}\label{s:epsilon}
%\emph{$\epsilon$-Core:}
%\\
When the \emph{core} set of a game is empty, it means no coalition
of players can gain anything by deviating. An outcome would be
unstable if a coalition can benefit even by a small amount from
deviating, which is a strong requirement. In fact, in some
situations, deviations can be costly, or players may have loyalty
to their coalitions, or even it can be computationally intractable
to find those small benefits. It would only make sense for a
coalition to deviate if the gain from a deviation exceeds the cost
of performing the deviation. \emph{$\epsilon$-core} relaxes the
notion of the core, and only requires that no coalition would
benefit significantly, or within a constant amount($\epsilon$) by
deviating (see Equation \ref{eq:core}).

@@

Coalition structure formation is the problem of finding the best
partition of web services into teams. In these settings, the
performance of an individual service is less important than the
\emph{social welfare} of the whole system, which is the sum of the
values of all teams. Having the game $(N,v)$, a coalition
structure $(CS)$ is \emph{socially optimal} if $CS$ belongs to set
$\operatorname*{arg\,max}_{CS} v(CS)$ where $v(CS)$ is the sum of
the values of all coalitions inside $CS$. $v(CS) = \sum_{C \in
CS}v(C)$.
%The outcome of a characteristic function game in coalition structure settings, consists of two parts; first a disjoint partition of players (agents) into coalitions, called a \emph{coalition structure} (CS) and second a \emph{payoff vector} as mentioned in cooperative game solution concepts, which distributes the value of each coalition among its members.

@@

In this section, we present different web service community models
and focus on the problem of how both web services and community
masters as rational entities would adopt strategies to maximize
their payoff.

\subsection {Web Services and One Community}

In this scenario, we assume the existence of a typical community
managed by its master web service and web services need to join it
to be able to get requests from the master. The community master
is characterized by a requests rate $(R_{M})$ from users. Each web
service comes with a given QoS ($QoS_{ws}$) from which the
throughput $T_{ws}$ is excluded. This exclusion allows us to build
our analysis on the particular value of $T_{ws}$. Throughput is
the average rate of tasks a web service can perform per time unit.
Therefore, the master web service is providing tasks with the rate
of $R_{ws}$ and web services perform tasks with average output
quality of $QoS_{ws}$ and a throughput rate of $T_{ws}$. In this
setting, we define the value of the coalition of the web services
within a community as a function increasing in both $T_{ws}$ and
$QoS_{ws}$ as follows:
\begin{equation*}
output(C) = \sum_{ws \in C}{(T_{ws} \times QoS_{ws})}
\end{equation*}
\setlength{\abovedisplayshortskip}{2pt}
\begin{equation}
v(C) = \left\{
  \begin{array}{l l}
    output(C) & \quad \text{if $\sum_{ws}{T_{ws}} \leq R_{M}$}\\
    output(C) \times \frac{R_{M}}{\sum_{ws}{T_{ws}}} & \quad \text{if $\sum_{ws}{T_{ws}} > R_{M}$}
  \end{array} \right.
\end{equation}

The output of a coalition of web services is a function of the
throughput and provided QoS. If the throughput rate is more than
the master's input request rate, it means the web services inside
the community are capable of serving more requests than the
demand. Considering this factor, the valuation function is
designed to balance the output performance so that it matches the
exact throughput rate and QoS the web service can provide within
the particular community.
%** In the case where the limited tasks are distributed among web services uniformly, the value of coalition would be the proportion of the average QoS times their throughput to rate of available requests. **

In this first scenario, we only consider one grand coalition and
analyze the system from the point of view of one single master web
service and a collection of web services. The master web service
decides which members can join
%or should leave%
the community and distributes the requests and income among its
community members (see Figure \ref{fig_sim1}).

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{s1.eps}`
\caption{Web Services and A Grand Community}
\label{fig_sim1}
\end{figure}

The membership decision is made based on throughput and \emph{QoS}
of the considered web service. The goal is to have quality web
services in the community so it stays stable and no other web
services would have incentives to deviate and leave the coalition
$C$. Therefore, a basic method would be to check the core of the
coalition $C$ considering all the current community members (all
web services already residing within the community) and the new
web service. This algorithm uses the \emph{Shapley value}
distribution method as described in Equation \ref{eq:shapley} to
distribute the gain of $v(C)$ among all the members and then
checks if the \emph{Shapley value} payoff vector for this
community having the characteristic function $v(C)$ is in the
\emph{core}. In the \emph{Shapley value} payoff vector, the payoff
for each web service $ws_i$ is calculated based on its marginal
contribution $v(C \cup {i}) - v(C)$ over all the possible
different permutations in which the coalition can be formed, which
makes the payoff distribution fair. Because of going through all
the possible permutations of subsets of $N$, the nature of the
\emph{Shapley value} is combinatorial. This makes it impractical
to use as the size of our coalitions grows. However, it is proven
that in convex games, the \emph{Shapley value} lies in the core
\cite{DBLP:conf/ijcai/GrecoMPS11, myerson1991game}. Thus, if the
\emph{Core} is non-empty, the payoff vector is a member of the
\emph{Core}. The following proposition is important to make our
algorithm tractable.
% so in our algorithm we check the core membership of this payoff vector.

@@

As a direct consequence of this proposition, in order to keep the
characteristic function convex, new web services should have more
marginal contribution as the coalition size grows.
